### 2022-12-05 definitions.tex
details

\begin{definition}
A {\em (resp. strongly) well-structured transition systems} (abbreviated as WSTS resp. S-WSTS) 
is a TS
$(S, \rightarrow, \leq)$
equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the wqo is (resp. strongly) compatible with the transition relation, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq t_1$  and $s_1 \rightarrow s_2$ , there exists 
	$t_2 \in S$ with 
	$s_2 \leq t_2$ and $t_1 \rightarrow^{*} t_2$ 			
				(resp. $t_1 \rightarrow^{1} t_2$ ).
\end{definition}

@@

\begin{proposition}
If $S$ is an WSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred^*(I )$ is upward-
closed.
\end{proposition}

@@

\begin{proposition}
If $S$ is an SWSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
\end{proposition}
\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.
\fi

\begin{definition}
A WSTS has {\em effective pred-basis} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
\end{definition}

@@

A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
$s \in J$. 
Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.

% \textcolor{red}{Do we still demand this ?}

@@

We ask whether we can reach a state 
%which satisfies
in 
%
$SAFE$  in a reasonable amount of time whenever we reach a state 
% which satisfies
in
%
$BAD$. 
From this we formulate two resilience problems. First consider the case where the recovery time
is bound by a given natural number $k \geq 0$, i.e., the explicit resilience problem for TS.

\problemx{$k$-resilience problem}
{a TS $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

We can also ask whether there exists such a bound $k$. We call this problem the resilience problem.

\problemx{Resilience problem}
{a TS $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\exists k \geq 0 ~ \forall s' \in BAD ~ (s \rightarrow^* s') \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

\begin{remark}
\begin{itemize}
\item expliquer toutes les variantes de la résilience
\item J=tout et I=$\uparrow x$ alors resilience = couverture.
\item J=tout et I=$\downarrow x$ alors resilience = accessibilité.
\end{itemize}
\end{remark}

@@

\alain{regarder les modèles aut à pile...}

@@

\cite{DBLP:conf/gg/Ozkan22} deduced that resilience is decidable for VASS because a finite basis of  $\uparrow post^*$ can be computed.

\begin{corollary}
{\sc resilience} is decidable for context-free grammars and VASS.
\end{corollary}

Context-free grammars and VASS are WSTS avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) is donc resilience calculable.

@@

\begin{proposition}
Given a WSTS with strong upward compatibility, an effective basis of $\uparrow post^*$ is not computable.
\end{proposition}

@@

\begin{proof}
Let $J=\downarrow Reach(S)$ and $I = \uparrow t$ where $t \in S$. Now resilience remains to the following property: $\forall s \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$. Let us translate this property for LCS for which Reach=$\downarrow Reach$: $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$. This property can be used to express the liveness property that there is an infinite run passing infinitely often through control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
\end{proof}

Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets but $Reach \neq \downarrow Reach$. Another proof could be a reduction of liveness to resilience and liveness is undecidable for many models like reset PN,...

@@

>>>>

\begin{definition}\cite{DBLP:journals/tcs/FinkelS01}
A {\em (resp. strongly) well-structured transition system} (abbreviated as WSTS)  $(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ between states such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the transition relation $ \rightarrow$ is compatible (resp. strongly compatible) with the wqo, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq t_1$  and $s_1 \rightarrow s_2$ , there exists 
	$t_2 \in S$ with 
	$s_2 \leq t_2$ and $t_1 \rightarrow^{*} t_2$ 			
				(resp. $t_1 \rightarrow t_2$ ).
\end{definition}

@@

\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
If $S$ is an WSTS and $I \subseteq S$ is an upward-closed set of states, then $\pred^*(I )$ is upward-
closed.
\end{proposition}

@@

\begin{proposition}\cite{DBLP:journals/tcs/FinkelS01}
If $S$ is a WSTS with strong monotony and $I \subseteq S$ is an upward-closed set of states, then $\pred(I )$ is upward-closed.
\end{proposition}
\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.
\fi

\begin{definition}\cite{DBLP:journals/iandc/AbdullaCJT00}
A WSTS has {\em effective pred-basis} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
\end{definition}

@@

A downard-closed set $J$ is {\em decidable} if, given $s \in S$, it is decidable whether
$s \in J$. 
Since a downward-closed set does not have an ``upward-basis'' in general, we will demand that membership is decidable.
\alain{remarque pas claire}
% \textcolor{red}{Do we still demand this ?}

@@

We ask whether we can reach a state 
%which satisfies
in 
%
$SAFE$  in a reasonable amount of time whenever we reach a state 
% which satisfies
in
%
$BAD$. 
From this we formulate two resilience problems. First consider the case where the recovery time
is bound by a given natural number $k \geq 0$, i.e., the \emph{explicit resilience problem} for TS.

\problemx{$k$-Resilience problem or Explicit Resilience problem}
{a transition system $(S,\rightarrow)$, an integer $k$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

If a system $S$ satisfies the explicit resilience property for an integer $k$, we say that $S$ is $k$-resilient.

@@

We can also ask whether there exists such a bound $k$ such that $S$ is $k$-resilient. We call this problem the \emph{bounded resilience problem}.

\problemx{Bounded Resilience problem}
{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\exists k \geq 0 ~ \forall s' \in BAD, ~ s \rightarrow^* s' \implies \exists s'' \in SAFE ~ s' \rightarrow^{\leq k} s''$ ?\newline}

\alain{repetition...} plus court: equivaut à demander si $\exists k \geq 0$ tel que $S$ est $k$-resilient.

\problemx{Bounded Resilience problem}
{a transition system $(S,\rightarrow)$, a state $s \in S$, $SAFE, BAD \subseteq S$ and $SAFE \cap BAD = \emptyset$.}
{$\exists k \geq 0  S$ is $k$-resilient. ?\newline}

pour $k,k'$ donnés, chercher si $SAFE_{max}$ et $BAD_{max}$ ont un sens ? on peut faire l'union des SAFE, SAFE' et des BAD, BAD' prendre le max des k. Fixer k, chercher et calculer $SAFE_{max}$ et $BAD_{max}$.

@@

\begin{proposition}
coverability, liveness and reachability reduce to resilience.
\end{proposition}

\begin{proof}

\begin{itemize}
\item J=tout et I=$\uparrow x$ alors resilience = couverture.
\item J=tout et I=$\downarrow x$ alors resilience = accessibilité.
\item J=tout sauf  exemple des LCS....
\end{itemize}
Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$. Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$. Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$. This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
\end{proof}

expliquer toutes les variantes de la résilience

@@

@@

\cite{DBLP:conf/gg/Ozkan22} deduced that resilience is decidable for VASS because VASS are a finite basis of  $\uparrow post^*$ can be computed.

\begin{corollary}
{\sc resilience} is decidable for context-free grammars.
\end{corollary}

Context-free grammars are WSTS avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) is donc resilience calculable.

@@

\begin{proposition}
Given a class of WSTS with strong upward compatibility, there don't exist an algorithm computing an effective basis of $\uparrow post^*$.
\end{proposition}

@@

\begin{proof} \alain{à relire/refaire}

\end{proof}

Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets. Another proof could be a reduction of liveness to resilience and since liveness is undecidable for many models like reset PN,...resilience is also undecidable for WSTS with strong monotony.

@@

\cite{DBLP:journals/ipl/BouajjaniEFMRWW00}
